---
title: "Simulation to Explore Inferences"
author: "mjurado"
date: "Thursday, March 12, 2015"
output: html_document
---

The goal of this analysis is to explore inferences in order to:

* compare the sample means versus the theoretical mean of the distribution
* compare the sample variance versus the theoretical variance of the distribution
* prove that the distribution is approximately normal

#### Comparing the sample mean with the theoretical mean
```{r}
# Given information

n <- 40 # number of exponentials
lambda <- .2 # rate
theo.mean <- round(1/ lambda) # <- theoretical mean of the distribution
std.dev <- theo.mean

# Calculate the sample mean
sample.mean <- round(mean(rexp(n,lambda)),3)

```
##### The theoretical mean is: `r theo.mean`
##### The sample mean is: `r sample.mean`

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
