---
title: 'Coursera Statistical Inference Project (Part 2): Basic Inferential Data Analysis'
author: "mjurado"
date: "Thursday, March 12, 2015"
output: pdf_document
---
***********************************************************************
###Overview
The goal of this analysis is to analyze the ToothGrowth dataset by summarizing the data as well as comparing tooth growth by supplement type and dosage.

***********************************************************************

####**Load the ToothGrowth data and perform some basic exploratory data analyses** 

```{r}
# Load Dataset
library(datasets)
data(ToothGrowth)

# structural summary
str(ToothGrowth)
# metric summary
summary(ToothGrowth)

# Comparison of Supplements and Dosage
table(ToothGrowth[,2:3])
```

####Provide a basic summary of the data.

* The data consists of a dataframe with 60 observations of 3 variables.
* The three variables are tooth length, supplement type, and dosage.
* The R help documentation states that the supplement type consists of two levels: Vitamin C (VC) and orange juice (OJ) and that the dosage is in milligrams.
* The min, max, and mean tooth lengths are: 4.20, 33.90, and 18.81, respectively.
* Half of the observations were given OJ and the other have Vitamin C.
* Each supplement type received three different dosages of 10 observations each

***********************************************************************

####**Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.** 

#####**1) Supplement Confidence Interval Testing**
```{r}
# Test the impact of supplement type on tooth growth
supp.test <- t.test(len ~ supp, data = ToothGrowth)
supp.test
```

**Supplement Results:**

The p-value is: `r round(supp.test$p.value,4)` which is greater than 5%. 

The 95% confidence interval is: (`r round(supp.test$conf.int,3)`) which contains 0.

As a result, we cannot reject the null hypothesis which means that the supplement type does not have a significant effect on increasing tooth length.

***********************************************************************

#####**2a) Dosage confidence interval testing: 0.5 - 1.0**
```{r}

tg.dose_0.5_1.0 <- subset(ToothGrowth, dose %in% c(0.5,1.0))
t.test_0.5_1.0 <- t.test(len ~ dose, data = tg.dose_0.5_1.0)
t.test_0.5_1.0


```

**Dosage 0.5-1.0 Results:** 

The p-value is: **`r t.test_0.5_1.0$p.value`** which is less than 5%. 

The 95% confidence interval is: **(`r round(t.test_0.5_1.0$conf.int,3)`)** which does not contain 0.

As a result, we can reject the null hypothesis.  This means that the application of doses between 0.5 and 1.0 has a significant effect on increasing tooth length.


#####**2b) Dosage confidence interval testing: 1.0 - 2.0**
```{r}

tg.dose_1.0_2.0 <- subset(ToothGrowth, dose %in% c(1.0,2.0))
t.test_1.0_2.0 <- t.test(len ~ dose, data = tg.dose_1.0_2.0)
t.test_1.0_2.0
```

**Dosage 1.0-2.0 Results:** 
The p-value is: **`r t.test_1.0_2.0$p.value`** which is less than 5%. 

The 95% confidence interval is: **(`r round(t.test_1.0_2.0$conf.int,3)`)** which does not contain 0.

As a result, we can reject the null hypothesis.  This means that the application of doses between 1.0 and 2.0 has a significant effect on increasing tooth length.


#####**2c) Dosage confidence interval testing: 0.5 - 2.0**
```{r}

tg.dose_0.5_2.0 <- subset(ToothGrowth, dose %in% c(0.5,2.0))
t.test_0.5_2.0 <- t.test(len ~ dose, data = tg.dose_0.5_2.0)
t.test_0.5_2.0
```

**Dosage 0.5-2.0 Results:**

The p-value is: **`r t.test_0.5_2.0$p.value`** which is less than 5%. 

The 95% confidence interval is: **(`r round(t.test_0.5_2.0$conf.int,3)`)** which does not contain 0.

As a result, we can reject the null hypothesis.  This means that the application of doses between 0.5 and 2.0 has a significant effect on increasing tooth length.

####*State your conclusions and the assumptions needed for your conclusion**

#####Conclusions:

* The type of supplement did not have a significant effect on tooth growth.
* 