---
title: 'Coursera Statistical Inference Project (Part 2): Basic Inferential Data Analysis'
author: "mjurado"
date: "Thursday, March 12, 2015"
output: html_document
---

####Overview
The goal of this analysis is to analyze the ToothGrowth dataset by summarizing the data as well as comparing tooth growth by supplement type and dosage.

####Analyses
#####Load the ToothGrowth data and perform some basic exploratory data analyses 

```{r}
# Load Dataset
library(datasets)
data(ToothGrowth)

# structural summary
str(ToothGrowth)
# metric summary
summary(ToothGrowth)

# Comparison of Supplements and Dosage
table(ToothGrowth[,2:3])
```

#####From the illustrations above, we can conclude that:

* The data consists of a dataframe with 60 observations of 3 variables.
* The three variables are tooth length, supplement type, and dosage.
* The R help documentation states that the supplement type consists of two levels: Vitamin C (VC) and orange juice (OJ) and that the dosage is in milligrams.
* The min, max, and mean tooth lengths are: 4.20, 33.90, and 18.81, respectively.
* Half of the observations were given OJ and the other have Vitamin C.
* Each supplement type received three different dosages of 10 observations each

```{r}
supp.mean = split(ToothGrowth$len,ToothGrowth$supp)
sapply(supp.mean, mean)

supp.mean = split(ToothGrowth$len,ToothGrowth$dose)
sapply(supp.mean, mean)
```



