---
title: 'Coursera Statistical Inference Project (Part 2): Basic Inferential Data Analysis'
author: "mjurado"
date: "Thursday, March 12, 2015"
output: pdf_document
geometry: margin=1.5cm
---
###Overview
The goal of this analysis is to analyze the effect on Vitamin C in guinea pigs using the ToothGrowth dataset from the R Dataset library.  The analysis will summarize the data as well as compare tooth growth by supplement type and dosage. 

***********************************************************************

###**Load the ToothGrowth data and perform some basic exploratory data analyses** 


```{r}
# Load Dataset
library(datasets)
data(ToothGrowth)

# structural summary
str(ToothGrowth)

# Comparison of Supplements and Dosage
table(ToothGrowth[,2:3])

```

###Provide a basic summary of the data.

* This data measured the effect on Vitamin C in guinea pigs and consisted of a dataframe with 60 observations of 3 variables
* The three variables are tooth length, supplement type, and dosage
* The R help documentation states that the supplement type consists of two levels: Vitamin C (VC) and orange juice (OJ) and that the dosage is in milligrams
* Half of the supplement type observations were given OJ and the other half were Vitamin C
* Each supplement type received three different dosages of 0.5, 1.0, and 2.0 for 10 observations each
* For a summary of the min, max, median and quartiles, see **Appendix A**



###**Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.** 

#####**1) Supplement Confidence Interval Testing**  


```{r}
# Test the impact of supplement type on tooth growth
supp.test <- t.test(len ~ supp, data = ToothGrowth)
supp.test$p.value
supp.test$conf.int
supp.test$estimate

```

**Supplement Results:**

The p-value is: $`r round(supp.test$p.value,4)`$ which is greater than 5%. 

The 95% confidence interval is: (`r round(supp.test$conf.int,3)`) which contains 0.


#####**2) Dosage confidence interval testing**  

```{r}
# Determine p-value, intervals, and means for 0.5 and 1.0
tg.dose_0.5_1.0 <- subset(ToothGrowth, dose %in% c(0.5,1.0))
t.test_0.5_1.0 <- t.test(len ~ dose, data = tg.dose_0.5_1.0)
t.test_0.5_1.0

# Determine p-value, intervals, and means for 1.0 and 2.0
tg.dose_1.0_2.0 <- subset(ToothGrowth, dose %in% c(1.0,2.0))
t.test_1.0_2.0 <- t.test(len ~ dose, data = tg.dose_1.0_2.0)
t.test_1.0_2.0

# Determine p-value, intervals, and means for 0.5 and 2.0
tg.dose_0.5_2.0 <- subset(ToothGrowth, dose %in% c(0.5,2.0))
t.test_0.5_2.0 <- t.test(len ~ dose, data = tg.dose_0.5_2.0)
t.test_0.5_2.0

```

**Dosage 0.5-1.0 Results:**  
The p-value is: **$`r t.test_0.5_1.0$p.value`$** which is less than 5%.  
The 95% confidence interval is: **($`r round(t.test_0.5_1.0$conf.int,3)`$)** which does not contain 0.

**Dosage 1.0-2.0 Results:**  
The p-value is: **$`r t.test_1.0_2.0$p.value`$** which is less than 5%.  
The 95% confidence interval is: **($`r round(t.test_1.0_2.0$conf.int,3)`$)** which does not contain 0.

**Dosage 0.5-2.0 Results:**  
The p-value is: **$`r t.test_0.5_2.0$p.value`$** which is less than 5%.  
The 95% confidence interval is: **($`r round(t.test_0.5_2.0$conf.int,3)`$)** which does not contain 0.

###**State your conclusions and the assumptions needed for your conclusion**  

#####Conclusions
* For the supplement testing, we cannot reject the null hypothesis, and therefore,we cannot conclude the supplement type has any significant effect on increasing tooth length
* For all three dosages, the p-values were less than 5% and none of the intervals contained 0.  Therefore we can reject the null hypothesis and conclude that dosage amounts have an impact on tooth growth
* In addition, it appears that the greater the dosage, the greater the impact on average growth

#####Assumptions
* All guinea pigs were randomly chosen to participate 
* All guinea pigs were of the same type, age, size
* All guinea pigs were given the same diet 
* All samples were unpaired with unequal variances
* The Welch approximation was used

\pagebreak

###APPENDIX A: Summary Of Min, Max, Median, and Quartiles 
```{r}
# Show summary of data (min,quartiles,median, max)
boxplot(len~supp*dose, data=ToothGrowth, notch=FALSE, 
        col=(c("navajowhite2","navajowhite3","navajowhite4")),
        main="Supplement Type/Dosage Observations", 
        xlab="Type.Dose (mg)", 
        ylab="Tooth Length") 
```